---
title: "NC_shelter_reimbursement_analysis"
author: "John Cornelison"
date: "12/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Research Notes

* There are instances when there is an operating and cost-per-animal cost with 0 intakes. I assume this is taking care of existing animals
* There are instances when there is no operating and cost-per-animal cost with multiple intakes. Just didn't get the operation cost.
* Is the cost-per-animal calculated by the number of animals entering the shelter that year, or the total number of animals in the shelter from previous year too. The cost per animal is not calculated by the current years intake. So, I feel like it is possible that higher cost could be an indicator of longer LOS. It also could be an indicator of type of shelter/rescue. There may be individual person rescues for example.

Live Release Rate: Adopted + Returned / Adopted + Returned + Euthanized

Boxplot cleaning info:

When removing years where facilities didn't report an operating cost, operation costs less than 6, and outliers, there were 1072 instances from 2010 to 2022




```{r, include = FALSE}


```


```{r, include = FALSE}
#Working Directory
#Desktop
setwd("C:/Users/john/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")

#Laptop
#setwd("C:/Users/johnc/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")


###Libraries
library(plotly)
library(patchwork)
library(tidyverse)
library(readxl)
library(googlesheets4)
library(stringr)
library(reactable)

###Reading in Data
NC_shelter_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Shelter Final")
NC_reimbursements_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Reimbursement Final")


```


```{r, include=FALSE}
#This block of code was used for sharing information with colaborator


# species_condensed_count <- species_table_condensed %>%
#   group_by(Species_Condensed) %>%
#   summarise(Count = sum(Intake))
# 
# 
# species_count <- species_table_condensed %>% 
#   group_by(Species_Condensed, Species) %>% 
#   summarise(Count = sum(Intake))


# gs4_deauth()
# gs4_auth()
# 
# my_sheet <- gs4_create("Animal Count and Catagorization")
# 
# write_sheet(species_condensed_count, ss = my_sheet)
# write_sheet(species_count, ss = my_sheet)
# 






```


```{r, include = FALSE}
#Data Cleaning

species_table <- NC_shelter_raw %>% 
  select(1:4, 7:874) %>% 
  pivot_longer(
    cols = 5:872,
    names_to = c("Species", "Category"),
    names_sep = "_",
    values_to = c("Count")) %>%
  pivot_wider(
    names_from = "Category",
    values_from = "Count") %>% 
  mutate(remove_row = case_when(Intake == 0 & Adopted == 0 & Euthanized == 0 & Returned == 0 & Euthanized == 0 ~ TRUE,
                                TRUE ~ FALSE)) %>% 
  filter(remove_row == FALSE) %>% 
  select(-remove_row) %>% 
  mutate(Intake = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 624,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 1012,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Intake),
         Adopted = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 380,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 98,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Adopted),
         Returned = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 58,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 16,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Returned),
         Euthanized = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 186,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 899,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Euthanized)) %>%
  #2020
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Cat", Intake = 387, Adopted = 213, Returned = 13, Euthanized = 77) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Raccoon", Intake = 10, Adopted = 0, Returned = 0, Euthanized = 3) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Opossum", Intake = 3, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Fox", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Bat", Intake = 4, Adopted = 0, Returned = 0, Euthanized = 2) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Skunk", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "71", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Species = "Dog", Intake = 385, Adopted = 207, Returned = 106, Euthanized = 193) %>%
  add_row(LIC = "71", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Species = "Cat", Intake = 628, Adopted = 324, Returned = 10, Euthanized = 362) %>% 
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Species = "Dog", Intake = 51, Adopted = 34, Returned = 9, Euthanized = 7) %>%
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Species = "Cat", Intake = 142, Adopted = 32, Returned = 1, Euthanized = 109) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Dog", Intake = 926, Adopted = 376, Returned = 359, Euthanized = 134) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Cat", Intake = 530, Adopted = 394, Returned = 33, Euthanized = 74) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Rabbit", Intake = 25, Adopted = 25, Returned = 0, Euthanized = 0) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Raccoon", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 10) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Opossum", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 3) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Bat", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 3) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Coyote", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 2) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Chicken", Intake = 15, Adopted = 8, Returned = 2, Euthanized = 5) %>% 
  #2021
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Dog", Intake = 451, Adopted = 60, Returned = 136, Euthanized = 59) %>%
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Cat", Intake = 494, Adopted = 142, Returned = 8, Euthanized = 125) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Rabbit", Intake = 17, Adopted = 2, Returned = 2, Euthanized = 6) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Raccoon", Intake = 17, Adopted = 2, Returned = 2, Euthanized = 7) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Opossum", Intake = 16, Adopted = 2, Returned = 1, Euthanized = 7) %>%
  add_row(LIC = "147", Year = 2021, "County/City" = "Martin", Facility_Name = "Martin County Sheriff'S Office Animal Shelter",
          Species = "Dog", Intake = 208, Adopted = 90, Returned = 42, Euthanized = 78) %>% 
  add_row(LIC = "147", Year = 2021, "County/City" = "Martin", Facility_Name = "Martin County Sheriff'S Office Animal Shelter",
          Species = "Cat", Intake = 280, Adopted = 65, Returned = 2, Euthanized = 200) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Dog", Intake = 1000, Adopted = 521, Returned = 233, Euthanized = 139) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Cat", Intake = 564, Adopted = 457, Returned = 26, Euthanized = 62) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Rabbit", Intake = 12, Adopted = 12, Returned = 1, Euthanized = 2) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Raccoon", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 30) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Opossum", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 22) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Bat", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 4) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Chicken", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 7) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Fox", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 4) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Turtle", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 2) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Squirrel", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 5) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Deer", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Bear", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Gerbil", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 6) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Alligator", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Coyote", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "115", Year = 2021, "County/City" = "Northampton", Facility_Name = "Northampton County Animal Shelter",
          Species = "Dog", Intake = 194, Adopted = 153, Returned = 30, Euthanized = 11) %>%
  add_row(LIC = "115", Year = 2021, "County/City" = "Northampton", Facility_Name = "Northampton County Animal Shelter",
          Species = "Cat", Intake = 188, Adopted = 167, Returned = 3, Euthanized = 18) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Dog", Intake = 437, Adopted = 42, Returned = 69, Euthanized = 67) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Cat", Intake = 632, Adopted = 86, Returned = 11, Euthanized = 431) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Rabbit", Intake = 3, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Raccoon", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Cattle", Intake = 26, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Duck", Intake = 4, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Chicken", Intake = 47, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Turkey", Intake = 9, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Species = "Goat", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>% 
  mutate(Outcome = Adopted + Euthanized + Returned) %>% 
  relocate(Outcome, .after = Intake)

#write_csv(species_table, "species_table.csv")

expenses_table <- NC_shelter_raw %>% 
  select(1:6) %>% 
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
          Operating_Expense = "316891", Cost_Per_Animal = "394.63") %>% 
  add_row(LIC = "72", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Operating_Expense = "475000", Cost_Per_Animal = "468.90") %>%
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Operating_Expense = "11214.67", Cost_Per_Animal = "57.81") %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Operating_Expense = "1537722", Cost_Per_Animal = "870") %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Operating_Expense = "1104472", Cost_Per_Animal = "1110.02") %>% 
  add_row(LIC = "147", Year = 2021, "County/City" = "Martin", Facility_Name = "Martin County Sheriff'S Office Animal Shelter",
           Operating_Expense = "119576.57", Cost_Per_Animal = "408.84") %>% 
  add_row(LIC = "16", Year = 2021, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
           Operating_Expense = "293568", Cost_Per_Animal = "890") %>% 
  add_row(LIC = "115", Year = 2021, "County/City" = "Northampton", Facility_Name = "Northampton County Animal Shelter",
          Operating_Expense = "48394.56", Cost_Per_Animal = "126.69") %>% 
  add_row(LIC = "63", Year = 2021, "County/City" = "Stokes", Facility_Name = "Stokes County Animal Shelter",
          Operating_Expense = "330320", Cost_Per_Animal = "290.70") %>% 
  group_by(Operating_Expense) %>% 
  mutate(Operating_Expense = as.numeric(Operating_Expense),
         Cost_Per_Animal = as.numeric(Cost_Per_Animal),
         LIC = as.numeric(LIC))

#write_csv(expenses_table, "expenses_table.csv")
  

reimbursements_table <- NC_reimbursements_raw %>% 
  mutate(across(c(5:28), as.numeric))

### Reading in Inflation adjusted expenses table and condensed species table

#Desktop
setwd("C:/Users/john/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")

#Laptop
#setwd("C:/Users/johnc/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")


species_table_condensed <- read_csv("species_table_condensed.csv") %>% 
  select(-...1) %>% 
  mutate(Species_Condensed = stringr::str_to_title(Species_Condensed),
         live_outcome = Adopted + Returned) 

NC_inflation_adjusted_expenses_table <- read_csv("inf_adj_expenses_table.csv") %>% 
  mutate(Year = as.character(Year))



```




There were 3 different outcomes listed for each shelter, adopted, returned, and euthanized. The adopted, returned, and euthanized rate is calculated by outcome / sum(3 outcomes).

```{r, echo = FALSE, warning=FALSE}
#Functions

cat_dog_other <- species_table_condensed %>% 
  mutate(cat_dog_other_col = case_when(Species_Condensed == "Cat" ~ "Cat",
                                       Species_Condensed == "Dog" ~ "Dog",
                                       TRUE ~ "Other")) %>% 
  group_by(Year, cat_dog_other_col) %>% 
  summarise(total_intakes = sum(Intake),
            total_live_outcomes = sum(live_outcome),
            total_euth = sum(Euthanized),
            LOR = total_live_outcomes/sum(Outcome),
            euth_rate = total_euth/sum(Outcome))
  

plot <- species_table_condensed %>% 
      filter(!Species_Condensed %in% c("Cat", "Dog")) %>% 
  group_by(Year, Species_Condensed) %>% 
  summarise(total_intakes = sum(Intake),
            total_live_outcomes = sum(live_outcome),
            total_euth = sum(Euthanized),
            LOR = total_live_outcomes/sum(Outcome),
            euth_rate = total_euth/sum(Outcome))


create_plotly_graph <- function(yvar, yaxis_title){
  
  plot <- species_table_condensed %>% 
    filter(!Species_Condensed %in% c("Cat", "Dog")) %>% 
  group_by(Year, Species_Condensed) %>% 
  summarise(total_intakes = sum(Intake),
            total_live_outcomes = sum(live_outcome),
            total_euth = sum(Euthanized),
            LOR = total_live_outcomes/sum(Outcome),
            euth_rate = total_euth/sum(Outcome)) %>%
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~yvar, color = ~Species_Condensed)
  
  plot <- plot %>% layout(yaxis = list(title = yaxis_title))
  
  
}


create_cat_dog_other_graph <- function(yvar, yaxis_title){
  
  plot <- cat_dog_other %>% 
  group_by(Year, cat_dog_other_col) %>% 
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~yvar, color = ~cat_dog_other_col)
  
  plot <- plot %>% layout(yaxis = list(title = yaxis_title))
  
  
}

#Use plot df to select yvars

```



```{r, echo = FALSE, warning= FALSE, include=FALSE}

intake_plot <- create_plotly_graph(plot$total_intakes, "Intakes")

LON_plot <- create_plotly_graph(plot$total_live_outcomes, "Number of Live Outcomes")

LOR_plot <- create_plotly_graph(plot$LOR, "Live Outcome Rate")

Euth_N_plot <- create_plotly_graph(plot$total_euth, "Number of Euthanizations")

Euth_R_plot <- create_plotly_graph(plot$euth_rate, "Euthanization Rate")

CDO_intake_plot <- create_cat_dog_other_graph(cat_dog_other$total_intakes, "Intakes")

CDO_LON_plot <- create_cat_dog_other_graph(cat_dog_other$total_live_outcomes, "Number of Live Outcomes")

CDO_LOR_plot <- create_cat_dog_other_graph(cat_dog_other$LOR, "Live Outcome Rate")

CDO_Euth_N_plot <- create_cat_dog_other_graph(cat_dog_other$total_euth, "Number of Euthanizations")

CDO_Euth_R_plot <- create_cat_dog_other_graph(cat_dog_other$euth_rate, "Euthanization Rate")




species_condensed_count <- species_table_condensed %>%
  group_by(Species_Condensed) %>%
  summarise(Count = sum(Intake))

intake_species_count_table <- reactable(species_condensed_count)


#cat_dog_intake_graph <- create_plotly_intake_graph(c("Cat", "Dog"))

#higher_intake_other_graph <- create_plotly_intake_graph(c("Raccoon", "Bird", "Pocket Pet", "Wildlife", "Rabbit"))

#lower_intake_other_graph <- create_plotly_intake_graph(c("Reptile", "Chicken", "Other", "Farm Animal", "Bat", "Opossum"))


aggregate_by_year <- species_table %>%
  mutate(Species = case_when(Species == "Dog" ~ "Dog",
                             Species == "Cat" ~ "Cat",
                             TRUE ~ "Other")) %>%
  group_by(Year, Species) %>% 
  summarise(total_intakes = sum(Intake),
            total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~total_intakes, color = ~Species)



outcome_rates_by_year <- species_table %>%
  filter(Species %in% c("Dog", "Cat", "Raccoon", "Rabbit", "Opossum")) %>% 
  group_by(Species, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  mutate(adopted_rate = round(total_adopted/total_outcomes, 2),
         returned_rate = round(total_returned/total_outcomes, 2),
         euthanized_rate = round(total_euthanized/total_outcomes, 2))

live_release_all <- species_table_condensed %>% 
  group_by(Species_Condensed, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_live_outcomes = sum(live_outcome)) %>% 
  mutate(live_release_rate = total_live_outcomes/total_outcomes)


cat_dogs_base_table <- species_table_condensed %>% 
  filter(Species_Condensed %in% c("Cat", "Dog")) %>% 
  group_by(Species_Condensed, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_live_outcomes = sum(live_outcome),
            total_euth = sum(Euthanized),
            total_intake = sum(Intake)) %>% 
  mutate(live_release_rate = total_live_outcomes/total_outcomes,
         euthanized_rate = total_euth/total_outcomes) 

live_release_other <- species_table_condensed %>% 
  filter(!Species_Condensed %in% c("Cat", "Dog")) %>% 
  group_by(Species_Condensed, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_live_outcomes = sum(live_outcome)) %>% 
  mutate(live_release_rate = total_live_outcomes/total_outcomes)


greater_66 <- live_release_other %>% 
  filter(Year == 2022 & live_release_rate > 0.66) %>%
  pull(Species_Condensed)

less_66 <- live_release_other %>% 
  filter(Year == 2022 & live_release_rate <= 0.66) %>%
  pull(Species_Condensed)


greater_66_dat <- live_release_other %>% 
  filter(Species_Condensed %in% greater_66)

less_66_dat <- live_release_other %>% 
  filter(Species_Condensed %in% less_66)


live_release_cat_dogs_graphic <- plot_ly(data = cat_dogs_base_table, type = "scatter", mode = "line", x = ~Year, y = ~live_release_rate, color = ~Species_Condensed)

intake_cat_dogs_plot <- plot_ly(data = cat_dogs_base_table, type = "scatter", mode = "line", x = ~Year, y = ~total_intake, color = ~Species_Condensed)

live_release_count_cats_dogs_plot <- plot_ly(data = cat_dogs_base_table, type = "scatter", mode = "line", x = ~Year, y = ~total_live_outcomes, color = ~Species_Condensed)

euth_count_cats_dogs_plot <- plot_ly(data = cat_dogs_base_table, type = "scatter", mode = "line", x = ~Year, y = ~total_euth, color = ~Species_Condensed)

euth_rate_cats_dogs_plot <- plot_ly(data = cat_dogs_base_table, type = "scatter", mode = "line", x = ~Year, y = ~euthanized_rate, color = ~Species_Condensed)


live_release_greater_other <- plot_ly(data = greater_66_dat, type = "scatter", mode = "line", x = ~Year, y = ~live_release_rate, color = ~Species_Condensed)

live_release_less_other <- plot_ly(data = less_66_dat, type = "scatter", mode = "line", x = ~Year, y = ~live_release_rate, color = ~Species_Condensed)

live_release_other_graphic <- plot_ly(data = live_release_other, type = "scatter", mode = "line", x = ~Year, y = ~live_release_rate, color = ~Species_Condensed)



adopted_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~adopted_rate, color = ~Species)

returned_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~returned_rate, color = ~Species)

euthanized_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~euthanized_rate, color = ~Species)








### Looking at opperating expense and cost_per_animal

op_cost <- NC_inflation_adjusted_expenses_table %>% 
  filter(is.na(Operating_Expense_inf_adj) == FALSE & Operating_Expense_inf_adj > 5)

Q1 <- quantile(op_cost$Operating_Expense_inf_adj, 0.25)
Q3 <- quantile(op_cost$Operating_Expense_inf_adj, 0.75)
IQR <- Q3 - Q1

# Define bounds
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

# Identify outliers
op_cost$outliers <- op_cost$Operating_Expense_inf_adj < lower_bound | op_cost$Operating_Expense_inf_adj > upper_bound


op_cost_boxplot <- op_cost %>% 
  filter(outliers == FALSE)  %>% 
  plot_ly(y = ~Operating_Expense_inf_adj, x = ~Year, type = "box")

# Customize the layout (optional)
op_cost_boxplot <- op_cost_boxplot %>% layout(
    title = "Operating Cost",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Cost")
)


```


## Do rising tides really lift all boats


#### Cat and Dog Plots

We can see from these plots that over time, the outcomes of cats and dogs have gotten better, but can we say the same for other animals.


```{r, echo = FALSE, warning = FALSE}

intake_cat_dogs_plot 

live_release_count_cats_dogs_plot 

live_release_cat_dogs_graphic

euth_count_cats_dogs_plot

euth_rate_cats_dogs_plot


```


#### Intakes Species Condensed Count Table

```{r, echo=FALSE, warning=FALSE}

intake_species_count_table

```

## Intake, Live Outcome, and Euthanized Graphics of Other Animals{.tabset}

I created two instances of the same graphics for the purpose of making comparisons.

### Intake 

```{r, echo =FALSE, warning=FALSE}

intake_plot

```

### LiveOutcomeN

```{r, echo = FALSE, warning=FALSE}

LON_plot

```


### LiveOutcomeR

```{r, echo = FALSE, warning=FALSE}

LOR_plot

```


### EuthN

```{r, echo = FALSE, warning=FALSE}

Euth_N_plot

```


### EuthR

```{r, echo = FALSE, warning=FALSE}

Euth_R_plot

```


## {-}


## {.tabset}

### Intake 

```{r, echo =FALSE, warning=FALSE}

intake_plot

```

### LiveOutcomeN

```{r, echo = FALSE, warning=FALSE}

LON_plot

```


### LiveOutcomeR

```{r, echo = FALSE, warning=FALSE}

LOR_plot

```


### EuthN

```{r, echo = FALSE, warning=FALSE}

Euth_N_plot

```


### EuthR

```{r, echo = FALSE, warning=FALSE}

Euth_R_plot

```


## {-}

## Cat, Dog, and Other Animals Graphics {.tabset}


### Intake 

```{r, echo =FALSE, warning=FALSE}

CDO_intake_plot

```

### LiveOutcomeN

```{r, echo = FALSE, warning=FALSE}

CDO_LON_plot

```


### LiveOutcomeR

```{r, echo = FALSE, warning=FALSE}

CDO_LOR_plot

```


### EuthN

```{r, echo = FALSE, warning=FALSE}

CDO_Euth_N_plot

```


### EuthR

```{r, echo = FALSE, warning=FALSE}

CDO_Euth_R_plot

```


## {-}

### John's Findings

* We can see that Intake has drastically decreased for Cats and Dogs. By about half for both cats and dogs. We can also see that the total number of live outcomes has remained about the same for Dogs between the years 2010 and 2019. Interestingly, cats have higher number of live outcomes. However, For both cats and dogs, the live release rate has increased by about 40%. With almost no increase in the number of live outcomes for dogs, we can probably assume that this increase in live release rate is heavily influenced by the lower number of intakes for dogs. We can also assume it played an impact in cats higher live release rate, but higher number of live outcomes also played a part.

* Compared to the number of cats and dogs, other animals make up a tiny portion of all animals entering the shelter. There are about 3.1 million dogs and cats that entered shelter within the last 12 years compared to only about 100,000 other animals. They made up only 3.13 percent of total animals.

* One of the primary goals of shelters is to find new homes for cats and dogs, but what about the other animals. If we look at the intake for Raccoons, Opossums, Birds, and Bats,  we can see that over the course of the 12 years, their intake has lowered, some at a higher rate than others. However, if we look at their euthanization rate, it has stayed about the same from 2010 to 2022 with dips and peaks between. The euthanization count has lowered, but this is most likely because there are less of these animals coming in.

* Each of the animals have a purpose/role in our society. Some are valuable, some are companions, and some are pests. Depending on which category the animal falls into, plays a big part in what happens to them. Lets take farm animals for example. If we look at their intake, we can see that both 2010 and 2022 had about the same number of farm animals entering the shelter with some peaks and valleys in between. We can also see that there live outcome ratio never dipped below 0.80 and got has high as 0.96. In part, this may because farm animals are valuable in that they provide food, or resources we can use. A farmer may be incentivized to pick up a low priced cow from the shelter if they have room to care for it.

* When looking at just Raccoon, Wildlife, and Bats, we can see that there live outcome ratio is less than 0.10 which in turn means their euthanization rate is extremely high. We can see that the euthanization trend line closely follows the intake trendline. Potentially, the reason so many raccoons and bats are euthanized is because of the potential diseases that they carry. Both bats and raccoons carry rabies. So, instead of spending resources to find a way to release the raccoons and bats in the wild, it is cheaper to simply euthanize them since they are not considered pets, can be percieved to be a nucience, and a public health hazard if released close to other people. It is best if people just leave them be instead of taking them to an animal shelter. Not exactly sure what the best solution is. It does not feel right just euthanizing them, but don't want anyone to contract a disease from them. I wonder how much it would cost to relocate them. For the wildlife category, I believe the high rate may be caused from euthanizing injured wildlife as well as not being able to relocate them to the wild. What do you do with a deer that wanders into someones yard? 


### Older Plots

#### Intake Graphs for Other Animals

```{r, warning=FALSE, echo=FALSE}

#higher_intake_other_graph

#live_release_other_graphic

#lower_intake_other_graph


```


#### Live Release Rate for Other Animals

We have 11 other animal groups. Since it made the graphic look a little crazy, the other animals were split between two graphics

We can see that some animals have higher live release rates than other animals. Compared to the dog and cat graphic, the increase for some other animals was much more sporadic. Sometimes jumping up by 0.5 on the live relase rate


```{r, echo = FALSE, warning = FALSE}

live_release_greater_other

live_release_less_other

live_release_other_graphic


```



#### Total Intakes by Year in North Carolina


```{r, echo = FALSE}
aggregate_by_year

```


#### Adoption Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

adopted_rate

```


#### Return Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

returned_rate


```


#### Euthanization Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums

```{r, echo = FALSE}


euthanized_rate


```

