---
title: "NC_shelter_reimbursement_analysis"
author: "John Cornelison"
date: "12/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Research Notes

* There are instances when there is an operating and cost-per-animal cost with 0 intakes. I assume this is taking care of existing animals
* There are instances when there is no operating and cost-per-animal cost with multiple intakes. Just didn't get the operation cost.
* Is the cost-per-animal calculated by the number of animals entering the shelter that year, or the total number of animals in the shelter from previous year too. The cost per animal is not calculated by the current years intake. So, I feel like it is possible that higher cost could be an indicator of longer LOS. It also could be an indicator of type of shelter/rescue. There may be individual person rescues for example.

Boxplot
When removing years where facilities didn't report an operating cost, operation costs less than 6, and outliers, there are 1072 instances from 2010 to 2022



```{r, include = FALSE}


```


```{r, include = FALSE}
#Working Directory
#Desktop
setwd("C:/Users/john/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")

#Laptop
#setwd("C:/Users/johnc/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")


###Libraries
library(plotly)
library(tidyverse)
library(readxl)
library(googlesheets4)
library(stringr)


###Reading in Data
NC_shelter_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Shelter Final")
NC_reimbursements_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Reimbursement Final")

species_table_condensed <- read_csv("species_table_condensed.csv") %>% 
  select(-...1) %>% 
  mutate(Species_Condensed = stringr::str_to_title(Species_Condensed))

NC_inflation_adjusted_expenses_table <- read_csv("inf_adj_expenses_table.csv") %>% 
  mutate(Year = as.character(Year))


```


```{r, include=FALSE}
#This block of code was used for sharing information with colaborator


species_condensed_count <- species_table_condensed %>%
  group_by(Species_Condensed) %>%
  summarise(Count = sum(Intake))
# 
# 
# species_count <- species_table_condensed %>% 
#   group_by(Species_Condensed, Species) %>% 
#   summarise(Count = sum(Intake))


# gs4_deauth()
# gs4_auth()
# 
# my_sheet <- gs4_create("Animal Count and Catagorization")
# 
# write_sheet(species_condensed_count, ss = my_sheet)
# write_sheet(species_count, ss = my_sheet)
# 






```


```{r, include = FALSE}
#Data Cleaning

species_table <- NC_shelter_raw %>% 
  select(1:4, 7:874) %>% 
  pivot_longer(
    cols = 5:872,
    names_to = c("Species", "Category"),
    names_sep = "_",
    values_to = c("Count")) %>%
  pivot_wider(
    names_from = "Category",
    values_from = "Count") %>% 
  mutate(remove_row = case_when(Intake == 0 & Adopted == 0 & Euthanized == 0 & Returned == 0 & Euthanized == 0 ~ TRUE,
                                TRUE ~ FALSE)) %>% 
  filter(remove_row == FALSE) %>% 
  select(-remove_row) %>% 
  mutate(Intake = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 624,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 1012,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Intake),
         Adopted = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 380,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 98,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Adopted),
         Returned = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 58,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 16,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Returned),
         Euthanized = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 186,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 899,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Euthanized)) %>%
  #2020
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Cat", Intake = 387, Adopted = 213, Returned = 13, Euthanized = 77) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Raccoon", Intake = 10, Adopted = 0, Returned = 0, Euthanized = 3) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Opossum", Intake = 3, Adopted = 0, Returned = 0, Euthanized = 1) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Fox", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Bat", Intake = 4, Adopted = 0, Returned = 0, Euthanized = 2) %>%
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
                             Species = "Skunk", Intake = 1, Adopted = 0, Returned = 0, Euthanized = 0) %>%
  add_row(LIC = "71", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Species = "Dog", Intake = 385, Adopted = 207, Returned = 106, Euthanized = 193) %>%
  add_row(LIC = "71", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Species = "Cat", Intake = 628, Adopted = 324, Returned = 10, Euthanized = 362) %>% 
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Species = "Dog", Intake = 51, Adopted = 34, Returned = 9, Euthanized = 7) %>%
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Species = "Cat", Intake = 142, Adopted = 32, Returned = 1, Euthanized = 109) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Dog", Intake = 926, Adopted = 376, Returned = 359, Euthanized = 134) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Cat", Intake = 530, Adopted = 394, Returned = 33, Euthanized = 74) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Rabbit", Intake = 25, Adopted = 25, Returned = 0, Euthanized = 0) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Raccoon", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 10) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Opossum", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 3) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Bat", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 3) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Coyote", Intake = 0, Adopted = 0, Returned = 0, Euthanized = 2) %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Species = "Chicken", Intake = 15, Adopted = 8, Returned = 2, Euthanized = 5) %>% 
  #2021
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Dog", Intake = 451, Adopted = 60, Returned = 136, Euthanized = 59) %>%
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Cat", Intake = 494, Adopted = 142, Returned = 8, Euthanized = 125) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Rabbit", Intake = 17, Adopted = 2, Returned = 2, Euthanized = 6) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Raccoon", Intake = 17, Adopted = 2, Returned = 2, Euthanized = 7) %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Species = "Opossum", Intake = 16, Adopted = 2, Returned = 1, Euthanized = 7) %>%
  
  mutate(Outcome = Adopted + Euthanized + Returned) %>% 
  relocate(Outcome, .after = Intake)

write_csv(species_table, "species_table.csv")

expenses_table <- NC_shelter_raw %>% 
  select(1:6) %>% 
  add_row(LIC = "92", Year = 2020, "County/City" = "Jackson", Facility_Name = "Jackson County Animal Shelter",
          Operating_Expense = "316891", Cost_Per_Animal = "394.63") %>% 
  add_row(LIC = "72", Year = 2020, "County/City" = "Stanly", Facility_Name = "Stanly County Animal Protective Ser",
          Operating_Expense = "475000", Cost_Per_Animal = "468.90") %>%
  add_row(LIC = "129", Year = 2020, "County/City" = "Tyrrell", Facility_Name = "Tyrrell County Animal Shelter",
          Operating_Expense = "11214.67", Cost_Per_Animal = "57.81") %>% 
  add_row(LIC = "16", Year = 2020, "County/City" = "New Hanover", Facility_Name = "New Hanover Co Sheriffs Ofc Animal Svcs Unit",
          Operating_Expense = "1537722", Cost_Per_Animal = "870") %>% 
  add_row(LIC = "377", Year = 2021, "County/City" = "Chatham", Facility_Name = "Chatham County Sheriff'S Office Animal Resource Center",
          Operating_Expense = "1104472", Cost_Per_Animal = "1110.02") %>% 
  group_by(Operating_Expense) %>% 
  mutate(Operating_Expense = as.numeric(Operating_Expense),
         Cost_Per_Animal = as.numeric(Cost_Per_Animal),
         LIC = as.numeric(LIC))
  
  



reimbursements_table <- NC_reimbursements_raw %>% 
  mutate(across(c(5:28), as.numeric))

## Some of the numbers needed to be updated because of revised late submissions

## 2019
species_table_prac <- species_table %>%
  select(-Outcome) %>% 
  mutate(Intake = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 624,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 1012,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Intake),
         Adopted = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 380,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 98,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Adopted),
         Returned = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 58,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 16,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 0,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 55,
                            TRUE ~ Returned),
         Euthanized = case_when(Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Dog" ~ 186,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Cat" ~ 899,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Raccoon" ~ 64,
                            Facility_Name == "Montgomery Co. Animal Control Facility" & Year == 2019 & Species == "Opossum" ~ 0,
                            TRUE ~ Euthanized)) %>%
  #2020
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  add_row(LIC = 92, Year = 2020, "County/City" = "Jackson", Facility_name = "Jackson County Animal Shelter",
                             Species = "Dog", Intake = 416, Adopted = 149, Returned = 79, Euthanized = 18) %>%
  mutate(Outcome = Adopted + Euthanized + Returned)


```


# Graphics

There were 3 different outcomes listed for each shelter, adopted, returned, and euthanized. The adopted, returned, and euthanized rate is calculated by outcome / sum(3 outcomes).


```{r, echo = FALSE, warning= FALSE, include=FALSE}

aggregate_by_year <- species_table %>%
  mutate(Species = case_when(Species == "Dog" ~ "Dog",
                             Species == "Cat" ~ "Cat",
                             TRUE ~ "Other")) %>%
  group_by(Year, Species) %>% 
  summarise(total_intakes = sum(Intake),
            total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~total_intakes, color = ~Species)



outcome_rates_by_year <- species_table %>%
  filter(Species %in% c("Dog", "Cat", "Raccoon", "Rabbit", "Opossum")) %>% 
  group_by(Species, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  mutate(adopted_rate = round(total_adopted/total_outcomes, 2),
         returned_rate = round(total_returned/total_outcomes, 2),
         euthanized_rate = round(total_euthanized/total_outcomes, 2))


adopted_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~adopted_rate, color = ~Species)

returned_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~returned_rate, color = ~Species)

euthanized_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~euthanized_rate, color = ~Species)



### Looking at opperating expense and cost_per_animal

op_cost <- NC_inflation_adjusted_expenses_table %>% 
  filter(is.na(Operating_Expense_inf_adj) == FALSE & Operating_Expense_inf_adj > 5)

Q1 <- quantile(op_cost$Operating_Expense_inf_adj, 0.25)
Q3 <- quantile(op_cost$Operating_Expense_inf_adj, 0.75)
IQR <- Q3 - Q1

# Define bounds
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

# Identify outliers
op_cost$outliers <- op_cost$Operating_Expense_inf_adj < lower_bound | op_cost$Operating_Expense_inf_adj > upper_bound


op_cost_boxplot <- op_cost %>% 
  filter(outliers == FALSE)  %>% 
  plot_ly(y = ~Operating_Expense_inf_adj, x = ~Year, type = "box")

# Customize the layout (optional)
op_cost_boxplot <- op_cost_boxplot %>% layout(
    title = "Operating Cost",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Cost")
)




```

## Do rising tides really lift all boats {.tabset}


### Total Intakes by Year in North Carolina


```{r, echo = FALSE}
aggregate_by_year

```


### Adoption Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

adopted_rate

```


### Return Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

returned_rate


```


### Euthanization Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums

```{r, echo = FALSE}


euthanized_rate


```


### Operation Cost Boxplot

```{r, echo = FALSE}


op_cost_boxplot


```
