---
title: "NC_shelter_reimbursement_analysis"
author: "John Cornelison"
date: "12/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Research Notes

* There are instances when there is an operating and cost-per-animal cost with 0 intakes. I assume this is taking care of existing animals
* There are instances when there is no operating and cost-per-animal cost with multiple intakes. Just didn't get the operation cost.
* Is the cost-per-animal calculated by the number of animals entering the shelter that year, or the total number of animals in the shelter from previous year too. The cost per animal is not calculated by the current years intake. So, I feel like it is possible that higher cost could be an indicator of longer LOS. It also could be an indicator of type of shelter/rescue. There may be individual person rescues for example.

Boxplot
When removing years where facilities didn't report an operating cost, operation costs less than 6, and outliers, there are 1072 instances from 2010 to 2022



```{r, include = FALSE}


```


```{r, include = FALSE}
#Working Directory
#Desktop
#setwd("C:/Users/john/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")

#Laptop
setwd("C:/Users/johnc/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")


###Libraries
library(plotly)
library(tidyverse)
library(readxl)
library(googlesheets4)


###Reading in Data
NC_shelter_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Shelter Final")
NC_reimbursements_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Reimbursement Final")

species_table_condensed <- read_csv("species_table_condensed.csv") %>% 
  select(-...1)

NC_inflation_adjusted_expenses_table <- read_csv("inf_adj_expenses_table.csv") %>% 
  mutate(Year = as.character(Year))


```


```{r}
gs4_deauth()
gs4_auth()

library(googlesheets4)
gs4_auth(
  scopes = "https://www.googleapis.com/auth/spreadsheets"
)

my_sheet <- gs4_create("Animal Count and Catagorization")

write_sheet(species_condensed_count, ss = my_sheet)


species_condensed_count <- species_table_condensed %>% 
  group_by(Species_Condensed, Species) %>% 
  summarise(Count = sum(Intake))


species_count <- species_table %>% 
  group_by(Species) %>% 
  summarise(Count = sum(Intake))





```


```{r, include = FALSE}
#Data Cleaning


species_table <- NC_shelter_raw %>% 
  select(1:4, 7:874) %>% 
  pivot_longer(
    cols = 5:872,
    names_to = c("Species", "Category"),
    names_sep = "_",
    values_to = c("Count")) %>%
  pivot_wider(
    names_from = "Category",
    values_from = "Count") %>% 
  mutate(remove_row = case_when(Intake == 0 & Adopted == 0 & Euthanized == 0 & Returned == 0 & Euthanized == 0 ~ TRUE,
                                TRUE ~ FALSE)) %>% 
  filter(remove_row == FALSE) %>% 
  mutate(Outcome = Adopted + Euthanized + Returned) %>% 
  select(-remove_row) %>% 
  relocate(Outcome, .after = Intake)

write_csv(species_table, "species_table.csv")

expenses_table <- NC_shelter_raw %>% 
  select(1:6) %>% 
  group_by(Operating_Expense) %>% 
  mutate(Operating_Expense = as.numeric(Operating_Expense),
         Cost_Per_Animal = as.numeric(Cost_Per_Animal),
         LIC = as.numeric(LIC)) 



reimbursements_table <- NC_reimbursements_raw %>% 
  mutate(across(c(5:28), as.numeric))


```


# Graphics

There were 3 different outcomes listed for each shelter, adopted, returned, and euthanized. The adopted, returned, and euthanized rate is calculated by outcome / sum(3 outcomes).


```{r, echo = FALSE, warning= FALSE, include=FALSE}

aggregate_by_year <- species_table %>%
  mutate(Species = case_when(Species == "Dog" ~ "Dog",
                             Species == "Cat" ~ "Cat",
                             TRUE ~ "Other")) %>%
  group_by(Year, Species) %>% 
  summarise(total_intakes = sum(Intake),
            total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~total_intakes, color = ~Species)



outcome_rates_by_year <- species_table %>%
  filter(Species %in% c("Dog", "Cat", "Raccoon", "Rabbit", "Opossum")) %>% 
  group_by(Species, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  mutate(adopted_rate = round(total_adopted/total_outcomes, 2),
         returned_rate = round(total_returned/total_outcomes, 2),
         euthanized_rate = round(total_euthanized/total_outcomes, 2))


adopted_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~adopted_rate, color = ~Species)

returned_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~returned_rate, color = ~Species)

euthanized_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~euthanized_rate, color = ~Species)



### Looking at opperating expense and cost_per_animal

op_cost <- NC_inflation_adjusted_expenses_table %>% 
  filter(is.na(Operating_Expense_inf_adj) == FALSE & Operating_Expense_inf_adj > 5)

Q1 <- quantile(op_cost$Operating_Expense_inf_adj, 0.25)
Q3 <- quantile(op_cost$Operating_Expense_inf_adj, 0.75)
IQR <- Q3 - Q1

# Define bounds
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

# Identify outliers
op_cost$outliers <- op_cost$Operating_Expense_inf_adj < lower_bound | op_cost$Operating_Expense_inf_adj > upper_bound


op_cost_boxplot <- op_cost %>% 
  filter(outliers == FALSE)  %>% 
  plot_ly(y = ~Operating_Expense_inf_adj, x = ~Year, type = "box")

# Customize the layout (optional)
op_cost_boxplot <- op_cost_boxplot %>% layout(
    title = "Operating Cost",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Cost")
)




```


### Total Intakes by Year in North Carolina


```{r, echo = FALSE}
aggregate_by_year

```


### Adoption Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

adopted_rate

```


### Return Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

returned_rate


```


### Euthanization Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums

```{r, echo = FALSE}


euthanized_rate


```


### Operation Cost Boxplot

```{r, echo = FALSE}


op_cost_boxplot


```
