---
title: "NC_shelter_reimbursement_analysis"
author: "John Cornelison"
date: "12/14/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r, include = FALSE}
#Working Directory
#Desktop
#setwd("C:/Users/john/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")

#Laptop
setwd("C:/Users/johnc/Documents/GitHub_Data/VCAP-NC-Shelter-Reimbursement-Project-Data")


###Libraries
library(plotly)
library(tidyverse)
library(readxl)


###Reading in Data
NC_shelter_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Shelter Final")
NC_reimbursements_raw <- read_xlsx("Shelter&ReimburseReformatted.xlsx", sheet = "Reimbursement Final")

```


```{r, include = FALSE}
#Data Cleaning


species_table <- NC_shelter_raw %>% 
  select(1:4, 7:874) %>% 
  pivot_longer(
    cols = 5:872,
    names_to = c("Species", "Category"),
    names_sep = "_",
    values_to = c("Count")) %>%
  pivot_wider(
    names_from = "Category",
    values_from = "Count") %>% 
  mutate(remove_row = case_when(Intake == 0 & Adopted == 0 & Euthanized == 0 & Returned == 0 & Euthanized == 0 ~ TRUE,
                                TRUE ~ FALSE)) %>% 
  filter(remove_row == FALSE) %>% 
  mutate(Outcome = Adopted + Euthanized + Returned) %>% 
  select(-remove_row) %>% 
  relocate(Outcome, .after = Intake)

expenses_table <- NC_shelter_raw %>% 
  select(1:6) %>% 
  group_by(Operating_Expense) %>% 
  mutate(Operating_Expense = as.numeric(Operating_Expense),
         Cost_Per_Animal = as.numeric(Cost_Per_Animal),
         LIC = as.numeric(LIC)) 


reimbursements_table <- NC_reimbursements_raw %>% 
  mutate(across(c(5:28), as.numeric))


```


```{r}
animal_count <- species_table %>% 
  group_by(Species) %>% 
  summarise(Count = sum(Intake)) 

sum(animal_count$Count)

```


# Graphics

There were 3 different outcomes listed for each shelter, adopted, returned, and euthanized. The adopted, returned, and euthanized rate is calculated by outcome / sum(3 outcomes).


```{r, echo = FALSE, warning= FALSE, include=FALSE}

aggregate_by_year <- species_table %>%
  mutate(Species = case_when(Species == "Dog" ~ "Dog",
                             Species == "Cat" ~ "Cat",
                             TRUE ~ "Other")) %>%
  group_by(Year, Species) %>% 
  summarise(total_intakes = sum(Intake),
            total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  plot_ly(type = 'scatter', mode = 'line', x = ~Year, y = ~total_intakes, color = ~Species)



outcome_rates_by_year <- species_table %>%
  filter(Species %in% c("Dog", "Cat", "Raccoon", "Rabbit", "Opossum")) %>% 
  group_by(Species, Year) %>% 
  summarise(total_outcomes = sum(Outcome),
            total_adopted = sum(Adopted),
            total_returned = sum(Returned),
            total_euthanized = sum(Euthanized)) %>% 
  mutate(adopted_rate = round(total_adopted/total_outcomes, 2),
         returned_rate = round(total_returned/total_outcomes, 2),
         euthanized_rate = round(total_euthanized/total_outcomes, 2))


adopted_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~adopted_rate, color = ~Species)

returned_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~returned_rate, color = ~Species)

euthanized_rate <- plot_ly(data = outcome_rates_by_year, type = 'scatter', mode = 'line', x = ~Year, y = ~euthanized_rate, color = ~Species)

```


### Total Intakes by Year in North Carolina


```{r, echo = FALSE}
aggregate_by_year

```


### Adoption Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

adopted_rate

```


### Return Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums


```{r, echo = FALSE}

returned_rate


```


### Euthanization Rate of Dogs, Cats, Raccoons, Rabbits, and Oposssums

```{r, echo = FALSE}


euthanized_rate


```

